%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Chapter: Acknowledgements
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Acknowledgements}
\label{chap:acknowledgements}

This document represents the work of many people who have contributed to the PMIx community.
Without the hard work and dedication of these people this document would not have been possible.
The sections below list some of the active participants and organizations in the various PMIx standard iterations.

%%%%%%%%%% Version 2.0
\section{Version 2.0}

The following list includes some of the active participants in the PMIx v2 standardization process.

\begin{itemize}
\item Ralph H. Castain, Annapurna Dasari, Christopher A. Holguin, Andrew Friedley, Michael Klemm and Terry Wilmarth
\item Joshua Hursey, David Solt, Alexander Eichenberger, Geoff Paulsen, and Sameh Sharkawi
\item Aurelien Bouteiller and George Bosilca
\item Artem Polyakov, Igor Ivanov and Boris Karasev
\item Gilles Gouaillardet
\item Michael A Raymond and Jim Stoffel
\item Dirk Schubert
\item Moe Jette
\item Takahiro Kawashima and Shinji Sumimoto
\item Howard Pritchard
\item David Beer
\item Brice Goglin
\item Geoffroy Vallee, Swen Boehm, Thomas Naughton and David Bernholdt
\item Adam Moody and Martin Schulz
\item Ryan Grant and Stephen Olivier
\item Michael Karo
\end{itemize}

The following institutions supported this effort through time and travel support for the people listed above.

\begin{itemize}
\item Intel Corporation
\item IBM, Inc.
\item University of Tennessee, Knoxville
\item The Exascale Computing Project, an initiative of the US Department of Energy
\item National Science Foundation
\item Mellanox, Inc.
\item Research Organization for Information Science and Technology
\item HPE Co.
\item Allinea (ARM)
\item SchedMD, Inc.
\item Fujitsu Limited
\item Los Alamos National Laboratory
\item Adaptive Solutions, Inc.
\item INRIA
\item Oak Ridge National Laboratory
\item Lawrence Livermore National Laboratory
\item Sandia National Laboratory
\item Altair
\end{itemize}


%%%%%%%%%% Version 1.0
\section{Version 1.0}

The following list includes some of the active participants in the PMIx v1 standardization process.

\begin{itemize}
\item Ralph H. Castain, Annapurna Dasari and Christopher A. Holguin
\item Joshua Hursey and David Solt
\item Aurelien Bouteiller and George Bosilca
\item Artem Polyakov, Elena Shipunova, Igor Ivanov, and Joshua Ladd
\item Gilles Gouaillardet
\item Gary Brown
\item Moe Jette
\end{itemize}

The following institutions supported this effort through time and travel support for the people listed above.

\begin{itemize}
\item Intel Corporation
\item IBM, Inc.
\item University of Tennessee, Knoxville
\item Mellanox, Inc.
\item Research Organization for Information Science and Technology
\item Adaptive Solutions, Inc.
\item SchedMD, Inc.
\end{itemize}
