%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Chapter: Data Access Operations
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Data Access and Sharing}
\label{chap:data_sharing}

In \ac{PMIx} key-value pairs are the primary way that information is shared 
between processes in the \ac{PMIx} universe.  
A key-value pair consists of a string key and a value data structure
which includes the type of the value's data.  
Each key is unique within the context or domain in which it exists.
The context or domain of a key in PMIx is called a data realm.  
%% The pmix_info_t type (See \ref{chap:introduction:portability:attributes}), 
%% is also built on key-values pairs.  
%% However, pmix_info_t is used for influencing API calls and is not used for sharing 
%% information between processes.  
This chapter describes how key-value pairs are made available and accessed by clients.

A key-value pair has a scope defined by its data realm.  
PMIx defines data realms for sessions, jobs, applications, nodes, processes and a special 
realm for publish/lookup information.   
These are explained more fully in chapters \ref{chap:api_rsvd_keys} and \ref{chap:pub}.

The source or provider of key-value pairs can be the host environment or a \ac{PMIx} client.  
The host environment can provide key-values pairs in any realm except for the publish/lookup realm.
%%% Can we have a separate effort to change this restriction on publish/lookup for consistency?
%%% Would this get into too many other issues?
PMIx clients can only provide key-values pairs within their own process realm and the publish/lookup realm.
The value of a key is not necessarily static and can be modified by the original provider of the key.

\ac{PMIx} classifies keys as either reserved or non-reserved.
Only keys provided by a host environment or \ac{PMIx} implementation can begin with \code{"pmix"} and are called reserved keys.  
Host environments and \ac{PMIx} implementations must only provide keys that begin with \code{"pmix"} to avoid conflicts with keys that are provided by clients.
Reserved keys are documented in Chapter \ref{chap:api_rsvd_keys}.
Each standardized reserved key has an associated macro that is documented along with its description.
Likewise clients can only provide keys that do not begin with \code{"pmix"} to avoid conflicting with keys provided by the host environment or \ac{PMIx} implementation.   
The string representation and the macro are equivalent methods to refer to a key.
Throughout this document, the macro is used rather than its string value to refer to reserved keys.
Clients cannot create reserved keys, nor can they modify the value of reserved keys.
The way reserved keys are accessed by clients 
is different than for non-reserved keys as explained in Chapter \ref{chap:api_rsvd_keys}.

Host environments are not required to provide all keys that are documented in the standard.  
However, when providing keys that are documented, they must adhere to the behavior defined by the standard for reserved keys.  
An implementation cannot provide a reserved key that behaves differently than what is in the standard.  
Implementations are encouraged to standardize any non-standardized, reserved keys that they provide to avoid conflicting with other implementations or efforts to standardize the same key.

Host environments are permitted to provide non-reserved keys and non-standardized reserved keys.  
However there are important caveats to providing either of these categories of keys 
that are explained in Section \ref{chap:data_sharing:non_rsvd_keys} and Chapter \ref{chap:api_rsvd_keys}.
% Would like to make this statement stronger, but there is text in chapter 6 currently that
% says it is allowed to use "custom-prefixed keys which may adhere to either the reserved or non-reserved
% retrieval rules.
Host environments should avoid providing reserved keys which are not standardized.  
If the reserved key is later standardized to behave in a way that is different from the behavior 
provided by the host environment, the host environment will not be compliant with the PMIx standard.
 
Reserved keys are set by the host environment and are made available at client initialization. 
Each key is associated with a data realm and the retrieval API call will indicate the realm being queried.
Reserved keys are used to access information about a client's execution context.
After initialization, the client can access reserved keys using the \refapi{PMIx_Get} API.
Common information that a client may access includes, but is not limited to,
the name of the host it is executing on, 
the number of peer processes in its namespace,
the number of peer processes in its application,
and its job rank.

%%% (Do we want to make any other statements about how and when values can change during 
%%%  the course of execution.
%%%  PMIX_EXIT_CODE for example, UNIVERSE_SIZE could change if more resources are added?)

Non-reserved keys are provided by clients.
A call to \refapi{PMIx_Put} will result in a key-value being associated with the calling client within 
its process data realm.
There is no mechanism to retract a key-value that has been made available by a client through \refapi{PMIx_Put}.
The key can be assigned a new value, but cannot be removed entirely.
Although key values can originate from different sources, they are always retrieved using the
\refapi{PMIx_Get} and \refapi{PMIx_Get_nb} \ac{API}.

The publish/lookup data realm is accessed through a separate set of \acp{API}.
\refapi{PMIx_Put} cannot add or modify key-values within the publish/lookup realm and 
\refapi{PMIx_Lookup} cannot access key-values outside the publish/lookup realm.
This data realm is described in detail is chapter \ref{chap:pub}.
Although \refapi{PMIx_Publish} and \refapi{PMIx_Lookup} are analogous to \refapi{PMIx_Put} and \refapi{PMIx_Get} in that both pairs 
of \acp{API} provide functionality for exposing and retreiving key-values, the semantics vary significantly.  
For example, \refapi{PMIx_Lookup} includes the ability to wait until a key is defined before returning.
\refapi{PMIx_Publish} can restrict the scope of who can access data to its own namespace, the host environment, the session it is executing in, all processes or even custom scopes, while
\refapi{PMIx_Put} can restrict the scope of who can access data to the node it is executing on, remote nodes,
all nodes or only itself.
The publish/lookup data realm is useful for advertising information that is not necessarily specific to one process to other processes in the \ac{PMIx} universe.
The process accessing this information does \emph{not} need to know the identity of the process that provided the data.

\ac{PMIx} does not provide a mechanism to asynchronously notify a process about the availability of key-value information once it is made available by another process.
However, the nonblocking accessor interfaces (e.g., \refapi{PMIx_Get_nb}, \refapi{PMIx_Lookup_nb}) may provide a degree of asynchronous notification on information availability.

Process related key-value exchanges allow a \ac{PMIx} process to share information specific to 
itself, and access information specific to one or more processes in the \ac{PMIx} universe.
These interactions occur within the process data realm.
The 'put/commit/get' exchange pattern is often used to exchange process related information.
Optionally, a 'put/commit/fence/get' exchange pattern adds the 'fence' synchronization 
(and possible collective exchange) for applications that desire it.
Commonly, these exchange patterns are used in a 
\declareterm{business card exchange} (a.k.a. \declareterm{modex exchange}) 
where one \ac{PMIx} client shares its connectivity information, then other \ac{PMIx} clients access that information to establish a connection with that client.
In some environments that support ``instant-on'' all connectivity information for \ac{PMIx} clients is stored in the job-level information at process creation time and is accessible to the clients without the need to perform any additional key-value exchange.

Keys can exist in multiple data realms, possibly with different values. 
\ac{PMIx} clients can access available information associated with any of the data realms.
For example, a client can access the number of nodes (\refattr{PMIX_NUM_NODES}) used by a 
session, job, or application.
Rather than having three different attributes, a single attribute is used but with the 
data realm context of the query specified as additional attributes.
Examples of these access patterns are presented in Section \ref{chap:api_rsvd_keys:getex}.

\ac{PMIx} clients can share key-value pairs associated with themselves by using the \refapi{PMIx_Put} function.
The \refapi{PMIx_Put} function automatically associates the key-value pair with the calling process, thus making it specific to that process.
A client may call \refapi{PMIx_Put} as many times as necessary and the data is not available to other processes until explicitly committed.
A client must call \refapi{PMIx_Commit} to make accessible all key-value pairs previously put by this process to all other processes in the \ac{PMIx} universe.
This put and commit pattern provides implementors the opportunity to make individual \refapi{PMIx_Put} calls efficient local operations, and then make the whole set of key-value pairs accessible in a single step.

\ac{PMIx} clients can access the key-value pairs associated with any process data realm 
in the \ac{PMIx} universe (including the calling process) by passing the specific process 
name of the target process to the \refapi{PMIx_Get} and \refapi{PMIx_Get_nb} functions.
The \ac{PMIx} server local to the calling process will retrieve that key-value pair from 
the \ac{PMIx} server associated with the target process.
Clients can also access session, job, application, node, and namespace level information 
by using the \refapi{PMIx_Get} and \refapi{PMIx_Get_nb} functions as shown in Section 
\ref{chap:api_rsvd_keys:getex}.
% and summarized in Table~\ref{tab:key-value-realms}.  (Add in next release?)
The completion semantics for \refapi{PMIx_Get} and \refapi{PMIx_Get_nb} differ 
depending on the type of key and its availability.  See Sections 
\ref{chap:data_sharing:retrules} and \ref{chap:api_rsvd_keys:retrules}.
For example, if a non-reserved key is not available, the \refapi{PMIx_Get} or \refapi{PMIx_Get_nb} call 
will not complete, by default, until that key-value pair becomes available.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Non-reserved keys}
\label{chap:data_sharing:non_rsvd_keys}

\emph{Non-reserved keys} are keys whose string representation begin with a
prefix other than \code{"pmix"}. Such keys are defined by a
client when information needs to be exchanged between processes, for example, 
where connection information is required and the host environment does not
support the \refterm{instant on} option or where the host environment does not
provide a required piece of data. Other than the prefix, there are no
restrictions on the use or content of non-reserved keys.

\ac{PMIx} provides support for two methods of exchanging non-reserved keys:

\begin{itemize}
    \item Global, collective exchange of the information prior to retrieval. This is accomplished by executing a barrier operation that includes collection and exchange of the data provided by each process such that each process has access to the full set of data from all participants once the operation has completed. \ac{PMIx} provides the \refapi{PMIx_Fence} function (or its non-blocking equivalent) for this purpose, accompanied by the \refattr{PMIX_COLLECT_DATA} qualifier.
    \item Direct, on-demand retrieval of the information. No barrier or global exchange is conducted in this case. Instead, information is retrieved from the host where that process is executing upon request - i.e., a call to \refapi{PMIx_Get} results in a data exchange with the \ac{PMIx} server on the remote host. Various caching strategies may be employed by the host environment and/or \ac{PMIx} implementation to reduce the number of retrievals. Note that this method requires that the host environment both know the location of the posting process and support direct information retrieval.
\end{itemize}

Both of the above methods are based on retrieval from a specific process -
i.e., the \refarg{proc} argument to \refapi{PMIx_Get} must include both the
namespace and the rank of the process that posted the information. However, in
some cases, non-reserved keys are provided on a globally unique basis and the
retrieving process has no knowledge of the identity of the process posting the
key. This is typically found in legacy applications (where the originating
process identifier is often embedded in the key itself) and in unstructured
applications that lack rank-related behavior. In these cases, the key remains
associated with the namespace of the process that posted it, but is retrieved
by use of the \refconst{PMIX_RANK_UNDEF} rank. In addition, the keys must be
globally exchanged prior to retrieval as there is no way for the host to
otherwise locate the source for the information.

Note that the retrieval rules for non-reserved keys (detailed in Section 
\ref{chap:data_sharing:retrules}) differ significantly from those used for reserved keys.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Posting Key/Value Pairs}
\label{chap:data_sharing:set}

\ac{PMIx} clients can post non-reserved key-value pairs associated with themselves by using \refapi{PMIx_Put}. Alternatively, \ac{PMIx} clients can cache both reserved and non-reserved key-value pairs accessible only by the caller via the \refapi{PMIx_Store_internal} \ac{API}.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\code{PMIx_Put}}
\declareapi{PMIx_Put}

%%%%
\summary

Stage a key/value pair in prepartion for being made accessible to processes.

%%%%
\format

\copySignature{PMIx_Put}{1.0}{
pmix_status_t \\
PMIx_Put(pmix_scope_t scope, \\
\hspace*{9\sigspace}const pmix_key_t key, \\
\hspace*{9\sigspace}pmix_value_t *val);
}

\begin{arglist}
\argin{scope}{Distribution scope of the provided value (handle)}
\argin{key}{key (\refstruct{pmix_key_t})}
\argin{value}{Reference to a \refstruct{pmix_value_t} structure (handle)}
\end{arglist}

\returnstart
\begin{itemize}
\item \refconst{PMIX_ERR_BAD_PARAM} indicating a reserved key is provided in the \refarg{key} argument.
\end{itemize}
\returnend

%%%%
\descr

Post a key-value pair for distribution. Depending upon the \ac{PMIx} implementation, the posted value may be locally cached in the client's \ac{PMIx} library until \refapi{PMIx_Commit} is called.

The provided \refarg{scope} determines the ability of other processes to access the posted data, as defined in \specrefstruct{pmix_scope_t}.
Specific implementations may support different scope values, but all implementations must support at least \refconst{PMIX_GLOBAL}.

The \refstruct{pmix_value_t} structure supports both string and binary values.
\ac{PMIx} implementations are required to support heterogeneous environments by properly converting binary values between host architectures, and will copy the provided \refarg{value} into internal memory prior to returning from \refapi{PMIx_Put}.

\adviceuserstart
Note that keys starting with a string of ``\code{pmix}'' must not be used in calls to \refapi{PMIx_Put}. Thus, applications should never use a defined ``PMIX'' attribute as the key in a call to \refapi{PMIx_Put}.
\adviceuserend


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Scope of Put Data}
\declarestruct{pmix_scope_t}
\label{api:nres:scope}

\versionMarker{1.0}
The \refstruct{pmix_scope_t} structure is a \code{uint8_t} type that defines the availability of data passed to \refapi{PMIx_Put}.
The following constants can be used to set a variable of the type \refstruct{pmix_scope_t}. All definitions were introduced in version 1 of the standard unless otherwise marked.

Specific implementations may support different scope values, but all implementations must support at least \refconst{PMIX_GLOBAL}.
If a specified scope value is not supported, then the \refapi{PMIx_Put} call must return \refconst{PMIX_ERR_NOT_SUPPORTED}.

\begin{constantdesc}
%
\declareconstitemvalue{PMIX_SCOPE_UNDEF}{0}
Undefined scope.
%
\declareconstitemvalue{PMIX_LOCAL}{1}
The data is intended only for other application processes on the same node.
Data marked in this way will not be included in data packages sent to remote requesters - i.e., it is only available to processes on the local node.
%
\declareconstitemvalue{PMIX_REMOTE}{2}
The data is intended solely for applications processes on remote nodes.
Data marked in this way will not be shared with other processes on the same node - i.e., it is only available to  processes on remote nodes.
%
\declareconstitemvalue{PMIX_GLOBAL}{3}
The data is to be shared with all other requesting processes, regardless of location.
%
\versionMarker{2.0}
\declareconstitemvalue{PMIX_INTERNAL}{4}
The data is intended solely for this process and is not shared with other processes.
%
\end{constantdesc}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\code{PMIx_Store_internal}}
\declareapi{PMIx_Store_internal}

%%%%
\summary

Store some data locally for retrieval by other areas of the process.

%%%%
\format

\copySignature{PMIx_Store_internal}{1.0}{
pmix_status_t \\
PMIx_Store_internal(const pmix_proc_t *proc, \\
\hspace*{20\sigspace}const pmix_key_t key, \\
\hspace*{20\sigspace}pmix_value_t *val);
}

\begin{arglist}
\argin{proc}{process reference (handle)}
\argin{key}{key to retrieve (string)}
\argin{val}{Value to store (handle)}
\end{arglist}

\returnstart
\begin{itemize}
\item \refconst{PMIX_ERR_BAD_PARAM} indicating a reserved key is provided in the \refarg{key} argument.
\end{itemize}
\returnend

%%%%
\descr

Store some data locally for retrieval by other areas of the process.
This is data that has only internal scope - it will never be posted externally. Typically used to cache data obtained by means outside of \ac{PMIx} so that it can be accessed by various areas of the process.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\code{PMIx_Commit}}
\declareapi{PMIx_Commit}

%%%%
\summary

Make available to other processes all key-value pairs previously staged via \refapi{PMIx_Put}.

%%%%
\format

\copySignature{PMIx_Commit}{1.0}{
pmix_status_t PMIx_Commit(void);
}

\returnsimple

%%%%
\descr

\ac{PMIx} implementations may choose to locally cache non-reserved keys prior to submitting them for distribution.  \refapi{PMIx_Commit} initiates the operation of making available previously staged key-value pairs to other processes.  Depending on the implementation, this may involve transmitting the entire collection of data posted by the process to a server.  \refapi{PMIx_Commit} is an asynchronous operation that will immediately return to the caller while the data is staged in the background.

\adviceuserstart
Users are advised to always include the call to \refapi{PMIx_Commit} in case the local implementation requires it. Note that posted data will not necessarily be circulated during \refapi{PMIx_Commit}. Availability of the data by other processes upon completion of \refapi{PMIx_Commit} therefore still relies upon the exchange mechanisms described at the beginning of this chapter.
\adviceuserend


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Retrieval rules for non-reserved keys}
\label{chap:data_sharing:retrules}

Since non-reserved keys cannot, by definition, have been provided by the host
environment, their retrieval follows significantly different rules than those
defined for reserved keys (as detailed in Section \ref{chap:api_rsvd_keys}).
\refapi{PMIx_Get} for a non-reserved key will obey the
following precedence search:

\begin{enumerate}
    \item If the \refattr{PMIX_GET_REFRESH_CACHE} attribute is given, then the
    request is first forwarded to the local \ac{PMIx} server which will then
    update the client's cache. Note that this may not, depending upon
    implementation details, result in any action.

    \item Check the local \ac{PMIx} client cache for the requested key - if not found and either the \refattr{PMIX_OPTIONAL} or \refattr{PMIX_GET_REFRESH_CACHE} attribute was given, the search will stop at this point and return the \refconst{PMIX_ERR_NOT_FOUND} status.

    \item Request the information from the local \ac{PMIx} server. The server
    will check its cache for the specified key. If the
    value still isn't found and the \refattr{PMIX_IMMEDIATE} attribute was
    given, then the library shall return the \refconst{PMIX_ERR_NOT_FOUND}
    error constant to the requester. Otherwise, the \ac{PMIx} server library
    will take one of the following actions:
    \begin{compactitemize}
        \item If the target process has a rank of \refconst{PMIX_RANK_UNDEF},
        then this indicates that the key being requested is globally unique
        and \emph{not} associated with a specific process. In this case, the
        server shall hold the request until either the data appears at the
        server or, if given, the \refattr{PMIX_TIMEOUT} is reached. In the
        latter case, the server will return the \refconst{PMIX_ERR_TIMEOUT}
        status. Note that the server may, depending on \ac{PMIx}
        implementation, never respond if the caller failed to specify a
        \refattr{PMIX_TIMEOUT} and the requested key fails to arrive at the
        server.

        \item If the target process is \emph{local} (i.e., attached to the
        same \ac{PMIx} server), then the server will hold the request until
        either the target process provides the data or, if given, the
        \refattr{PMIX_TIMEOUT} is reached. In the latter case, the server will
        return the \refconst{PMIX_ERR_TIMEOUT} status. Note that data which is
        posted via \refapi{PMIx_Put} but not staged with \refapi{PMIx_Commit}
        may, depending upon implementation, never appear at the server.

        \item If the target process is \emph{remote} (i.e., not attached to
        the same \ac{PMIx} server), the server will either:
        \begin{compactitemize}
            \item If the host has provided the
            \refapi{pmix_server_dmodex_req_fn_t} module function
            interface, then the server
            shall pass the request to its host for servicing. The host is
            responsible for determining the location of the target process and
            passing the request to the \ac{PMIx} server at that location.

            When the remote data request is received, the target \ac{PMIx}
            server will check its cache for the specified key. If the key is
            not present, the request shall be held until either the target
            process provides the data or, if given, the \refattr{PMIX_TIMEOUT}
            is reached. In the latter case, the server will return the
            \refconst{PMIX_ERR_TIMEOUT} status. The host shall convey the
            result back to the originating \ac{PMIx} server, which will reply
            to the requesting client with the result of the request when the
            host provides it.

            Note that the target server may, depending on \ac{PMIx}
            implementation, never respond if the caller failed to specify a
            \refattr{PMIX_TIMEOUT} and the target process fails to post the
            requested key.

            \item if the host does not support the
            \refapi{pmix_server_dmodex_req_fn_t} interface, then
            the server will immediately respond to the client with the
            \refconst{PMIX_ERR_NOT_FOUND} status
        \end{compactitemize}
    \end{compactitemize}
\end{enumerate}

\adviceimplstart
While there is no requirement that all \ac{PMIx} implementations follow the
client-server paradigm used in the above description, implementers are
required to provide behaviors consistent with the described search pattern.
\adviceimplend

\adviceuserstart
Users are advised to always specify the \refattr{PMIX_TIMEOUT} value when
retrieving non-reserved keys to avoid potential deadlocks should the specified
key not become available.
\adviceuserend

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\code{PMIx_Get}}
\declareapi{PMIx_Get}

%%%%
\summary

Retrieve a key/value pair from the client's namespace.

%%%%
\format

\copySignature{PMIx_Get}{1.0}{
pmix_status_t \\
PMIx_Get(const pmix_proc_t *proc, const pmix_key_t key, \\
\hspace*{9\sigspace}const pmix_info_t info[], size_t ninfo, \\
\hspace*{9\sigspace}pmix_value_t **val);
}

\begin{arglist}
\argin{proc}{Process identifier - a \code{NULL} value may be used in place of the caller's ID (handle)}
\argin{key}{Key to retrieve (\refstruct{pmix_key_t})}
\argin{info}{Array of info structures (array of handles)}
\argin{ninfo}{Number of elements in the \refarg{info} array (integer)}
\argout{val}{value (handle)}
\end{arglist}

A successful return indicates that the requested data has been returned in the manner requested (.e.g., in a provided static memory location ).

\returnstart
\begin{itemize}
\item \refconst{PMIX_ERR_BAD_PARAM} A bad parameter was passed to the function call - e.g., the request included the \refattr{PMIX_GET_STATIC_VALUES} directive, but the provided storage location was \code{NULL}
\item \refconst{PMIX_ERR_EXISTS_OUTSIDE_SCOPE} The requested key exists, but was posted in a \emph{scope} (see Section \ref{api:nres:scope}) that does not include the requester.
\item \refconst{PMIX_ERR_NOT_FOUND} The requested data was not available.
\end{itemize}
\returnend

\reqattrstart
The following attributes are required to be supported by all \ac{PMIx} libraries:

\pasteAttributeItem{PMIX_OPTIONAL}
\pasteAttributeItem{PMIX_IMMEDIATE}
\pasteAttributeItem{PMIX_DATA_SCOPE}
\pasteAttributeItem{PMIX_SESSION_INFO}
\pasteAttributeItem{PMIX_JOB_INFO}
\pasteAttributeItem{PMIX_APP_INFO}
\pasteAttributeItem{PMIX_NODE_INFO}
\pasteAttributeItem{PMIX_GET_STATIC_VALUES}
\pasteAttributeItem{PMIX_GET_POINTER_VALUES}
\pasteAttributeItem{PMIX_GET_REFRESH_CACHE}

\reqattrend

\optattrstart
The following attributes are optional for host environments:

\pasteAttributeItem{PMIX_TIMEOUT}

\optattrend

%%%%
\descr

Retrieve information for the specified \refarg{key} associated with the process identified in the given \refstruct{pmix_proc_t}. See Chapters \ref{chap:api_rsvd_keys} and \ref{chap:data_sharing:non_rsvd_keys} for details on rules governing retrieval of information. Information will be returned according to provided directives:

\begin{itemize}
    \item In the absence of any directive, the returned \refstruct{pmix_value_t} shall be an allocated memory object. The caller is responsible for releasing the object when done.
    \item If \refattr{PMIX_GET_POINTER_VALUES} is given, then the function shall return a pointer to a \refstruct{pmix_value_t} in the \ac{PMIx} library's memory that contains the requested information.
    \item If \refattr{PMIX_GET_STATIC_VALUES} is given, then the function shall return the information in the provided \refstruct{pmix_value_t} pointer. In this case, the caller must provide storage for the structure and pass the pointer to that storage in the \refarg{val} parameter.  If the implementation cannot return a static value, then the call to \refapi{PMIx_Get} must return the \refconst{PMIX_ERR_NOT_SUPPORTED} status.
\end{itemize}

This is a blocking operation - the caller will block until the retrieval rules of Section \ref{chap:api_rsvd_keys:retrules} or \ref{chap:data_sharing:retrules} are met.

The \refarg{info} array is used to pass user directives regarding the get operation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\code{PMIx_Get_nb}}
\declareapi{PMIx_Get_nb}

%%%%
\summary

Nonblocking \refapi{PMIx_Get} operation.

%%%%
\format

\versionMarker{1.0}
\cspecificstart
\begin{codepar}
pmix_status_t
PMIx_Get_nb(const pmix_proc_t *proc, const pmix_key_t key,
            const pmix_info_t info[], size_t ninfo,
            pmix_value_cbfunc_t cbfunc, void *cbdata);
\end{codepar}
\cspecificend

\begin{arglist}
\argin{proc}{Process identifier - a \code{NULL} value may be used in place of the caller's ID (handle)}
\argin{key}{Key to retrieve (string)}
\argin{info}{Array of info structures (array of handles)}
\argin{ninfo}{Number of elements in the \refarg{info} array (integer)}
\argin{cbfunc}{Callback function (function reference)}
\argin{cbdata}{Data to be passed to the callback function (memory reference)}
\end{arglist}

\returnsimplenb

If executed, the status returned in the provided callback function will be one of the following constants:

\begin{itemize}
\item \refconst{PMIX_SUCCESS} The requested data has been returned.
\item \refconst{PMIX_ERR_EXISTS_OUTSIDE_SCOPE} The requested key exists, but was posted in a \emph{scope} (see Section \ref{api:nres:scope}) that does not include the requester.
\item \refconst{PMIX_ERR_NOT_FOUND} The requested data was not available.
\item a non-zero \ac{PMIx} error constant indicating a reason for the request's failure.
\end{itemize}

\reqattrstart
The following attributes are required to be supported by all \ac{PMIx} libraries:

\pasteAttributeItem{PMIX_OPTIONAL}
\pasteAttributeItem{PMIX_IMMEDIATE}
\pasteAttributeItem{PMIX_DATA_SCOPE}
\pasteAttributeItem{PMIX_SESSION_INFO}
\pasteAttributeItem{PMIX_JOB_INFO}
\pasteAttributeItem{PMIX_APP_INFO}
\pasteAttributeItem{PMIX_NODE_INFO}
\pasteAttributeItem{PMIX_GET_POINTER_VALUES}
\pasteAttributeItem{PMIX_GET_REFRESH_CACHE}

\divider

The following attributes are required for host environments that support this operation:

\pasteAttributeItem{PMIX_WAIT}

\reqattrend

\optattrstart
The following attributes are optional for host environments that support this operation:

\pasteAttributeItem{PMIX_TIMEOUT}

\optattrend

%%%%
\descr

The callback function will be executed once the retrieval rules of Chapters \ref{chap:api_rsvd_keys} or \ref{chap:data_sharing:retrules} are met.
See \refapi{PMIx_Get} for a full description. Note that the non-blocking form of this function cannot support the \refattr{PMIX_GET_STATIC_VALUES} attribute as the user cannot pass in the required pointer to storage for the result.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Retrieval attributes}
\label{chap:api_kg:attr}

The following attributes are defined for use by retrieval \acp{API}:

%
\declareAttribute{PMIX_OPTIONAL}{"pmix.optional"}{bool}{
Look only in the client's local data store for the requested value - do not request data from the \ac{PMIx} server if not found.
}
%
\declareAttribute{PMIX_IMMEDIATE}{"pmix.immediate"}{bool}{
Specified operation should immediately return an error from the \ac{PMIx} server if the requested data cannot be found - do not request it from the host \ac{RM}.
}
%
\declareAttribute{PMIX_GET_POINTER_VALUES}{"pmix.get.pntrs"}{bool}{
Request that any pointers in the returned value point directly to values in the key-value store. The user \emph{must not} release any returned data pointers.
}
%
\declareAttribute{PMIX_GET_STATIC_VALUES}{"pmix.get.static"}{bool}{
Request that the data be returned in the provided storage location. The caller is responsible for destructing the \refstruct{pmix_value_t} using the \refmacro{PMIX_VALUE_DESTRUCT} macro when done.
}
%
\declareAttribute{PMIX_GET_REFRESH_CACHE}{"pmix.get.refresh"}{bool}{
When retrieving data for a remote process, refresh the existing local data cache for the process in case new values have been put and committed by the process since the last refresh. Local process information is assumed to be automatically updated upon posting by the process. A \code{NULL} key will cause all values associated with the process to be refreshed - otherwise, only the indicated key will be updated. A process rank of \refconst{PMIX_RANK_WILDCARD} can be used to update job-related information in dynamic environments. The user is responsible for subsequently updating refreshed values they may have cached in their own local memory.
}
%
\declareAttribute{PMIX_DATA_SCOPE}{"pmix.scope"}{pmix_scope_t}{
Scope of the data to be searched in a \refapi{PMIx_Get} call.
}
%
\declareAttribute{PMIX_TIMEOUT}{"pmix.timeout"}{int}{
Time in seconds before the specified operation should time out (zero indicating infinite) and return the \refconst{PMIX_ERR_TIMEOUT} error.
Care should be taken to avoid race conditions caused by multiple layers (client, server, and host) simultaneously timing the operation.
}
%
\declareAttribute{PMIX_WAIT}{"pmix.wait"}{int}{
Caller requests that the \ac{PMIx} server wait until at least the specified number of values are found (a value of zero indicates \emph{all} and is the default).
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
